---
title: "homework_2"
author: "Xianchen Wang"
date: "2023-10-11"
output:
  pdf_document: default
  html_document: default
---

## Data loading

```{r}
library(sf)
library(here)
library(tidyverse)
library(janitor)
library(dplyr)
library(tmap)

WC <- st_read(here::here("prac_2",
              "Data",
              "Washington_Counties_with_Natural_Shoreline___washsh_area",
              "Washington_Counties_with_Natural_Shoreline___washsh_area.shp"))
#plot(WC)
#summary(WC)

SchoolData <- read_csv(here::here("prac_2",
                       "Data",
                       "Report_Card_Assessment_Data_2018-19_School_Year.csv"),
                       locale = locale(encoding = "latin1"),
                       na = c("NULL", "None"))
#class(SchoolData)
```

## Data cleaning

```{r}
Datatypelist <- SchoolData %>%
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")
#Datatypelist
#summary(SchoolData)

SchoolData <- SchoolData %>%
  clean_names()
#summary(SchoolData)

```

## Data manipulate

First method, calculate each row average passing rate, then calculate WC average passing rate, finally group by.

$$ \frac{A+B}{C+D} \neq \frac{A}{C} + \frac{B}{D} $$

```{r}
SchoolData_1 <- SchoolData %>%
  replace_na(list(count_met_standard = 0, count_of_students_expected_to_test = 1))
#slice_head(SchoolData, n=60)

Passing_rate_row_group_average <- SchoolData_1 %>%
  mutate(row_averagepassingrate = count_met_standard/
                                  count_of_students_expected_to_test) %>%
  group_by(county) %>%
  summarise(count=n(), averagepassingrate = sum(row_averagepassingrate)) %>%
  mutate(WC_passing_rate = mean(averagepassingrate)) %>%
  mutate(rate_compare = case_when(averagepassingrate - WC_passing_rate > 0 ~ "above", TRUE ~ "below"))
  
  #mutate(disparity= averagepassingrate - WC_passing_rate) %>%
  #dplyr::select(county, averagepassingrate, disparity, WC_passing_rate) #%>%
  #arrange(disparity)
```

## Mapping_1

```{r}
gradesMap_1 <- WC %>%
  clean_names() %>%
  left_join(.,
            Passing_rate_row_group_average,
            by = c("countylabe" = "county"))
  
tmap_mode("plot")
# map_1 <- qtm(gradesMap_1, fill = "rate_compare")
# tmap_save(map_1, 
#            "D:/CASA/CASA_0005/practical/prac_2/Map/Passing_rate_row_group_average_map.jpeg", 
#            dpi=600)
map_1 <- tm_shape(gradesMap_1)+
  tm_borders() +
  tm_fill(col = 'rate_compare',
          palette = c("Above" = "blue", "Below" = "red"),
          title = "Above or Below") +
  tm_compass(position = c("left", "bottom"),type = "arrow") +
  tm_layout(title = "WC_1",
            legend.position = c("right", "bottom"))
tmap_save(map_1, here::here("prac_2","Map","Passing_rate_row_group_average_map.jpeg"), dpi=600)

```

## Data manipulate

Second method, group by first then use average passing rate calculating WC passing rate.

$$ \frac{A+B}{C+D} \neq \frac{\frac{A}{C} + \frac{B}{D}}{2} $$

```{r}
SchoolData_2 <- SchoolData %>%
  replace_na(list(count_met_standard = 0, count_of_students_expected_to_test = 0))

Passing_rate_group_average <- SchoolData_2 %>%
  group_by(county) %>%
  summarise(count=n(), count_met_standard_all = sum(count_met_standard),
            count_all = sum(count_of_students_expected_to_test)) %>%
  mutate(averagepassingrate = count_met_standard_all / count_all) %>%
  mutate(WC_passing_rate = mean(averagepassingrate)) %>%
  mutate(rate_compare = case_when(averagepassingrate - WC_passing_rate > 0 ~ "above", TRUE ~ "below"))

```

## Mapping_2

```{r}
gradesMap_2 <- WC %>%
  clean_names() %>%
  left_join(.,
            Passing_rate_group_average,
            by = c("countylabe" = "county"))

tmap_mode("plot")
# tm_shape(gradesMap_2) +
#   tm_borders() +
#   tm_fill(col = "rate_compare",
#           palette = c("Above" = "blue", "Below" = "red"),
#           title = "Above or Below")
map_2 <- tm_shape(gradesMap_2)+
  tm_borders() +
  tm_fill(col = 'rate_compare',
          palette = c("Above" = "blue", "Below" = "red"),
          title = "Above or Below") +
  tm_compass(position = c("left", "bottom"),type = "arrow") +
  tm_layout(title = "WC_2",
            legend.position = c("right", "bottom"))
tmap_save(map_2, here::here("prac_2","Map","Passing_rate_group_average_map.jpeg"), dpi=600)
```

## Data manipulate

Third method, group by first then use all count and all met standard count calculating WC passing rate.

Is correct!

```{r}
SchoolData_3 <- SchoolData %>%
  filter(., test_subject == "Science") %>%
  filter(., grade_level == "All Grades") %>%
  filter(., count_of_students_expected_to_test != "NA") %>%
  filter(., county != "Multiple") %>%
  filter(., organization_level == "School")

# SchoolData_3 <- SchoolData %>%
#   replace_na(list(count_met_standard = 0, count_of_students_expected_to_test = 0))

Passing_rate_group_count <- SchoolData_3 %>%
  group_by(county) %>%
  summarise(count=n(), count_met_standard_all = sum(count_met_standard),
            count_all = sum(count_of_students_expected_to_test)) %>%
  mutate(averagepassingrate = count_met_standard_all / count_all) %>%
  # mutate(WC_passing_rate = mean(averagepassingrate)) %>%
  mutate(WC_passing_rate = sum(count_met_standard_all) / sum(count_all)) %>%
  mutate(rate_compare = case_when(averagepassingrate - WC_passing_rate > 0 ~ "above", TRUE ~ "below"))

```

## Mapping_3

```{r}
gradesMap_3 <- WC %>%
  clean_names() %>%
  left_join(.,
            Passing_rate_group_count,
            by = c("countylabe" = "county"))

tmap_mode("plot")
 tm_shape(gradesMap_3) +
   tm_borders() +
   tm_fill(col = "rate_compare",
           palette = c("Above" = "blue", "Below" = "red"),
           title = "Above or Below")
 map_3 <- tm_shape(gradesMap_3) +
   tm_borders() +
   tm_fill(col = 'rate_compare',
           palette = c("Above" = "blue", "Below" = "red"),
           title = "Above or Below") +
   tm_compass(position = c("left", "bottom"),type = "arrow") +
   tm_layout(title = "WC_3",
             legend.position = c("right", "bottom"))
# qtm(gradesMap_3, fill='rate_compare')
# ggplot(data = gradesMap_3) +
#   geom_sf(aes(fill = rate_compare), # 将income映射到填充颜色上
#           color = "black",  # 轮廓颜色
#           size = .5) + # 轮廓粗细
#   theme_minimal()
# ggsave(here::here("prac_2","Map","Passing_rate_group_count_map.jpeg"))
#tmap_save(map_3, "D:/CASA/CASA_0005/practical/prac_2/Map/1Passing_rate_group_count_map.jpeg")
tmap_save(map_3, here::here("prac_2","Map","Passing_rate_group_count_map.jpeg"))
```
